# Toy Data Wrangling Project

###### This project is based on a toy data set showing product purchases from an electronics store. There are four brands: Philips, Akzo, Van Houten, and Unilever. However, there are many different spellings and capitalizations of those names. Also, the product code and number are combined in one column, separated by a hypen. R is used to clean this data set to make it easier to visualize and analyze.

### load dplyr and tidyr and csv file
```{r load dplyr and tidyr and csv file}
library(dplyr)
library(tidyr)

refine_original <- read.csv("refine_original.csv")
```

### standardize company names
```{r standardize company names}
refine_company <- refine_original[,]

refine_company$company <- sub(".+s$", "philips", refine_company$company, ignore.case = TRUE)

refine_company$company <- sub(".+o$|.+0$", "akzo", refine_company$company, ignore.case = TRUE)

refine_company$company <- sub("^van houten$", "van houten", refine_company$company, ignore.case = TRUE)

refine_company$company <- sub("^unil.+r$", "unilever", refine_company$company, ignore.case = TRUE)
```

### separate product code and number
```{r separate product code and number}
refine_product <- refine_company[,]
colnames(refine_product)[2] <- "product_code_num"

refine_product <- separate(refine_product, product_code_num, c("product_code", "product_number"), sep="-")
```

### add product categories
```{r add product categories}
product_code <- c("p","v","x","q")
product_category <- c("Smartphone","TV","Laptop","Tablet")
product_categories <- data.frame(product_code, product_category)

refine_categories <- merge(refine_product, product_categories, by="product_code")

refine_categories <- refine_categories[, c(2,1,8,3,4,5,6,7)]
```

### add full address for geocoding
```{r add full address for geocoding}
refine_address <- refine_categories[,]

refine_address <- unite(refine_address, "full_address", c("address", "city", "country"), sep=",")
```

### create binary variables for company and product category
```{r create binary variables for company and product category}
refine_clean <- refine_address[,]

refine_clean$company_philips <- ifelse(refine_clean$company == "philips", 1, 0)

refine_clean$company_akzo <- ifelse(refine_clean$company == "akzo", 1, 0)

refine_clean$company_van_houten <- ifelse(refine_clean$company == "van houten", 1, 0)

refine_clean$company_unilever <- ifelse(refine_clean$company == "unilever", 1, 0)

refine_clean$product_smartphone <- ifelse(refine_clean$product_category == "Smartphone", 1, 0)

refine_clean$product_tv <- ifelse(refine_clean$product_category == "TV", 1, 0)

refine_clean$product_laptop <- ifelse(refine_clean$product_category == "Laptop", 1, 0)

refine_clean$product_tablet <- ifelse(refine_clean$product_category == "Tablet", 1, 0)
```

### write cleaned data to csv file
```{r write cleaned data to csv file}
write.csv(refine_clean, file="refine_clean.csv")
```